---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = NULL
)
```

# arkhe

<!-- badges: start -->
[![Appveyor build status](https://ci.appveyor.com/api/projects/status/7xt157mqcny0pht1?svg=true)](https://ci.appveyor.com/project/nfrerebeau/arkhe)
[![R build status](https://github.com/nfrerebeau/arkhe/workflows/R-CMD-check/badge.svg)](https://github.com/nfrerebeau/arkhe/actions)
[![codecov](https://codecov.io/gh/nfrerebeau/arkhe/branch/master/graph/badge.svg)](https://codecov.io/gh/nfrerebeau/arkhe)

[![CRAN Version](http://www.r-pkg.org/badges/version/arkhe)](https://cran.r-project.org/package=arkhe)
[![CRAN checks](https://cranchecks.info/badges/worst/arkhe)](https://cran.r-project.org/web/checks/check_results_arkhe.html)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/arkhe)](https://cran.r-project.org/package=arkhe)

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3526659.svg)](https://doi.org/10.5281/zenodo.3526659)
<!-- badges: end -->

## Overview

A collection of classes that represent archaeological data. This package provides a set of S4 classes that represent different special types of matrix (absolute/relative frequency, presence/absence data, co-occurrence matrix, etc.) upon which package developers can build subclasses. It also provides a set of generic methods (mutators and coercion mechanisms) and functions (e.g. predicates). In addition, a few classes of general interest (e.g. that represent stratigraphic relationships) are implemented.

## Installation

You can install the released version of **arkhe** from [CRAN](https://CRAN.R-project.org) with:

```{r cran-installation, eval=FALSE}
install.packages("arkhe")
```

Or install the development version from GitHub with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("nfrerebeau/arkhe")
```

## Usage

```{r packages}
# Load the package
library(arkhe)
```

```{r vignette, eval=FALSE}
# See the vignette
utils::vignette("arkhe", package = "arkhe")
```

**arkhe** provides a set of S4 classes that represent different special types of matrix.

* Numeric matrix:
    * `CountMatrix` represents absolute frequency data,
    * `AbundanceMatrix` represents relative frequency data,
    * `OccurrenceMatrix` represents a co-occurrence matrix,
    * `SimilarityMatrix` represents a (dis)similarity matrix,
* Logical matrix:
    * `IncidenceMatrix` represents presence/absence data,
    * `StratigraphicMatrix` represents stratigraphic relationships.

*It assumes that you keep your data tidy*: each variable (type/taxa) must be saved in its own column and each observation (assemblage/sample) must be saved in its own row.

These new classes are of simple use, on the same way as the base `matrix`:

```{r create}
# Define a count data matrix
quanti <- CountMatrix(data = sample(0:10, 100, TRUE), nrow = 10, ncol = 10)

# Define a logical matrix
# Data will be coerced with as.logical()
quali <- IncidenceMatrix(data = sample(0:1, 100, TRUE), nrow = 10, ncol = 10)
```

**arkhe** uses coercing mechanisms (with validation methods) for data type conversions:

```{r coerce}
## Create a count matrix
A0 <- matrix(data = sample(0:10, 100, TRUE), nrow = 10, ncol = 10)

## Coerce to absolute frequencies
A1 <- as_count(A0)

## Coerce to relative frequencies
B <- as_abundance(A1)

## Row sums are internally stored before coercing to a frequency matrix
## (use get_totals() to get these values)
## This allows to restore the source data
A2 <- as_count(B)
all(A1 == A2)

## Coerce to presence/absence
C <- as_incidence(A1)

## Coerce to a co-occurrence matrix
D <- as_occurrence(A1)
```

Represent stratigraphic relationships:

```{r stratigraphy}
# Principles of Archaeological Stratigraphy, fig. 12
harris <- read.table(
  header = TRUE,
  text = "lower upper
          2     1
          3     1
          4     1
          5     2
          5     3
          5     4
          6     5
          7     1
          7     6
          8     1
          8     6
          9     7
          9     8"
)

S <- as_stratigraphy(harris)
```

## Contributing
Please note that the **arkhe** project is released with a [Contributor Code of Conduct](https://github.com/nfrerebeau/arkhe/blob/master/.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
